---
title: "STA365Homework4"
author: "Stephen Shannon"
date: "March 29, 2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
library(rstan)


survey <- readr::read_csv(file = "survey.csv")
poststrat <- readr::read_csv(file = "poststrat.csv")
head(survey)
head(poststrat)
```


```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

Question 1) Write a stan program that fits a multilevel logisitic regression model to the survey data.

```{stan, output.var ="cat_pref_model"}
  data {
  int<lower = 0> n_survey;
  int<lower = 0> n_age;
  int<lower = 0> n_eth;
  int<lower = 0> n_income;
  int<lower = 0> n_state;
  int cat_pref[n_survey];
  vector[n_survey] male;
  int age[n_survey];
  int eth[n_survey];
  int income[n_survey];
  int state[n_survey];
  // Bit for poststratification
  int<lower = 0> n_pred;
  vector[n_pred] male_pred;
  int age_pred[n_pred];
  int eth_pred[n_pred];
  int N_in_cell_pred[n_pred];
}
parameters {
  real mu;
  real beta;
  vector[n_age] z_age;
  vector[n_eth] z_eth;
  vector[n_income] z_income;
  vector[n_state] z_state;
  real<lower= 0> tau_age;
  real<lower= 0> tau_eth;
  real<lower= 0> tau_income;
  real<lower= 0> tau_state;
}
transformed parameters {
  vector[n_age] alpha_age = tau_age * z_age;
  vector[n_eth] alpha_eth = tau_eth * z_eth;
  vector[n_income] alpha_income = tau_income * z_income;
  vector[n_state] alpha_state = tau_state * z_state;
}
model {
  cat_pref ~ binomial_logit(1, mu + beta*male + alpha_age[age] + alpha_eth[eth] + alpha_income[income] + alpha_state[state]);
  z_age ~ normal(0,1);
  z_eth ~ normal(0,1);
  z_state ~ normal(0,1);
  z_income ~ normal(0,1);
  tau_age ~ normal(0,1);
  tau_eth ~ normal(0,1);
  mu ~ normal(0,1);
  beta ~ normal(0,1);
}
generated quantities {
  int cat_pref_pred[n_pred];
  for (n in 1:n_pred) {
    cat_pref_pred[n] = binomial_rng(N_in_cell_pred[n], 
   1.0/(1.0 + exp(-(mu + beta*male_pred[n] + alpha_age[age_pred[n]] + alpha_eth[eth_pred[n]]))));
  }
}

```

Question 2) Use the poststratification matrix and the samples frm the posterior to get the posterior of the total proportion of cat lovers. Plot the posterior
```{r}
stan_data <- list(
  n_survey = length(survey$cat_pref),
  n_age = length(unique(poststrat$age)),
  n_eth = length(unique(poststrat$eth)),
  n_income = length(unique(poststrat$income)),
  n_state = length(unique(poststrat$state)),
  cat_pref = survey$cat_pref,
  male = survey$male,
  age = survey$age,
  eth = survey$eth,
  income = survey$income,
  state = survey$state,
  n_pred = length(poststrat$male),
  male_pred = poststrat$male,
  age_pred = poststrat$age,
  eth_pred = poststrat$eth,
  income_pred = poststrat$income,
  state_pred = poststrat$state,
  N_in_cell_pred = poststrat$N
)
fit <- sampling(cat_pref_model, data = stan_data, control=list(adapt_delta=0.99))

## Propotion of people saying cat_pref
n_pop <- sum(poststrat$N)
cat_pref <- rstan::extract(fit,"cat_pref_pred")
# This gives a 4000x6300 matrix so each row
# is a sample from the posterior predictive!
prop <- rowSums(cat_pref$cat_pref_pred)/n_pop
hist(prop,breaks=30)
```